---
title: "Milan active-travel transformations (2016–2021): Tract-level indicators and interactive maps from official and custom data"
format:
  html
---

## Overview

This report documents active-travel transformations in Milan (2016–2021) using official municipal datasets from the Comune di Milano and a custom dataset of *Piazze Aperte*.

In addition to interactive city-level maps, the report constructs **tract-level infrastructure indicators**, enabling subsequent linkage with electoral data.

## Data sources

- **Official municipal data (Comune di Milano)**
  - **Cycling infrastructure** (lines): *ciclabile sede propria*, *ciclabile segnaletica*
  - **Pedestrian-priority infrastructure** (polygons): *DT (divieto di transito)*, *AP (area pedonale)*

- **Custom dataset (this study)**
  - ***Piazze Aperte*** (polygons): manually digitised public-space interventions  
    Only interventions flagged for analysis were retained (**31 features**), and classified as:
    + Permanent
    + Tactical

## Data processing

### Completion of missing periods using GSV (pedestrian-priority areas)

For pedestrian-priority polygons, implementation timing was first extracted from the official field `val_inizio` when available.  
Some features had missing or unusable dates.

These cases were reviewed manually using **Google Street View (GSV)**, and an approximate **implementation period** was assigned based on visual evidence.  
The resulting period labels were then re-imported and joined back to the spatial dataset.

For features where a year was available but no manual period was assigned, period attribution followed the year-based rules described below.

## Period attribution (alignment with election cycles)

To enable a consistent temporal comparison aligned with political mandates in Milan, interventions were classified into broad **implementation periods based on municipal election cycles**.

Municipal elections took place in **June 2016** and **October 2021**. Because implementation dates are not consistently available at month-level precision, it is not possible to apply exact pre-/post-election cut-offs.  
We therefore use **election years as anchor points**, interpreting periods as approximations of electoral cycles.

The following periods were defined:

- `pre2016`: interventions dated before 2016  
- `2016–2021`: interventions dated between 2016 and 2021 (inclusive)  
- `post2021`: interventions dated after 2021 (not used in the main analysis)

### Justification of cut-offs

- **Why `pre2016` (not `pre2017`):**  
  The 2016 election occurred in June, making 2016 a transition year. Because month-level timing is not consistently available, we cannot reliably separate pre- and post-election interventions within 2016.  
  We therefore treat 2016 as the start of the new cycle and define `pre2016` as interventions dated **before 2016**.

- **Why `2016–2021` (not `2016–2022`):**  
  The next election took place in October 2021, and interventions dated **2022 onwards** clearly belong to the subsequent cycle.  
  Given the lack of precise timing, we treat **2022+ as `post2021`**, avoiding contamination of the 2016–2021 period.

These cut-offs provide a consistent framework for comparing infrastructure implementation across political cycles without implying unavailable temporal precision.

```{r}
#| message: false
#| warning: false
#| include: false

source("R/infra_00_setup.R")
source("R/infra_01_data_processing.R")
source("R/infra_02_tract_indicators.R")
source("R/infra_03_linear_maps.R")
source("R/infra_04_thematic_maps.R")
```

## Interactive city-level maps

Interactive maps support validation and exploratory analysis at both the city and tract level.

For cycling, pedestrian-priority infrastructure, and *Piazze Aperte*, three maps were produced:

1. Infrastructure implemented **before 2016**
2. Infrastructure implemented **up to 2021** (`pre2016 + 2016–2021`)
3. **Changes between periods**, highlighting interventions implemented during `2016–2021`

Census tract boundaries for Milan (created by Gabriele Pinto; labelled as “precincts” in the original dataset)
Maps also include layer toggles by infrastructure type and direct GSV links for visual validation.

```{r}
#| echo: false
#| message: false
#| warning: false

official_dashboard_3maps
```

## Construction of tract-level indicators

Infrastructure was aggregated to census tracts to enable linkage with electoral data.

Cycling infrastructure (lines) and *Piazze Aperte* (polygons) were combined with tracts boundaries to compute tract-level indicators.

To capture **local exposure to place-based interventions**, a **50 m buffer** was applied to both cycling infrastructure and *Piazze Aperte*. Interventions are assigned to a tract if their 50 m buffer intersects the tract, allowing nearby areas to be considered affected even when the intervention does not fall strictly within administrative boundaries.

The tract-level dataset includes both **continuous** and **binary** indicators. Binary indicators record whether a tract was exposed to an intervention. Continuous indicators capture the magnitude of the intervention. These include both absolute measures (e.g. metres of infrastructure) and relative measures normalised by the total street-network length within the tract.

The resulting dataset is stored in `off_prec_summ` and contains indicators for all implementation periods. However, the analysis focuses on **infrastructure implemented between 2016 and 2021**, which captures variation in exposure prior to the electoral outcome.

The main indicators used in the analysis are:

- `road_m`: total street-network length in the tract  

### Cycling infrastructure (50 m exposure, 2016–2021)

- `cycle_buf50_any_2016_2021`  
  Binary indicator equal to 1 if the tract lies within 50 m of any cycling infrastructure  

- `cycle_buf50_m_2016_2021`  
  Metres of **original cycling infrastructure** assigned to the tract under the 50 m rule  

- `cycle_buf50_share_network_2016_2021`  
  Share of the tract’s street network represented by cycling infrastructure assigned under the 50 m rule  

### *Piazze Aperte* (50 m exposure, 2016–2021)

- `piazze_buf50_any_2016_2021`  
  Binary indicator equal to 1 if the tract lies within 50 m of any *Piazze Aperte* intervention  

- `piazze_buf50_m_2016_2021`  
  Metres of **original street network inside the *Piazze Aperte* intervention** assigned to the tract under the 50 m rule  

- `piazze_buf50_share_network_2016_2021`  
  Share of the tract’s street network represented by *Piazze Aperte*-affected street metres assigned under the 50 m rule  

@tbl-1 presents an illustrative preview of the tract-level dataset.

```{r}
#| echo: false
#| label: tbl-1
#| message: false
#| warning: false

off_prec_summ |>
  sf::st_drop_geometry() |>
  dplyr::select(dplyr::any_of(c(
    "tract_id",
    "road_m",

    "piazze_buf50_any_2016_2021",
    "piazze_buf50_m_2016_2021",
    "piazze_buf50_share_network_2016_2021",

    "cycle_buf50_any_2016_2021",
    "cycle_buf50_m_2016_2021",
    "cycle_buf50_share_network_2016_2021"

  ))) |>
  dplyr::mutate(
    dplyr::across(where(is.numeric), ~ round(.x, 2))
  ) |>
  head(10) |>
  DT::datatable(
    caption = "Preview of tract-level exposure indicators (50 m buffer, 2016–2021)",
    options = list(
      scrollX = TRUE,
      pageLength = 10
    ),
    rownames = FALSE
  )
```


## Thematic maps (tract level)

The maps show complementary indicators of active-travel transformations (2016–2021), capturing different dimensions of exposure to interventions.

For both cycling infrastructure and *Piazze Aperte*, three indicators are presented:

- **Binary exposure (any):** whether a tract is within 50 m of an intervention  
- **Absolute exposure (metres):** magnitude of the original intervention assigned to the tract under the 50 m rule  
- **Relative exposure (share of street network):** exposure normalised by the total street-network length within the tract  

Together, these indicators capture:

- the **spatial reach** of interventions (binary exposure)  
- their **intensity** (metres of intervention)  
- and their **relative importance within the local network** (share of street network)  

### Cycling exposure: binary

```{r}
#| echo: false
cycle_any_thematic_leaflet
```

### Cycling exposure: metres of bike lane

```{r}
#| echo: false
cycle_m_thematic_leaflet
```

### Cycling exposure: share of street network

```{r}
#| echo: false
cycle_share_thematic_leaflet
```

### Piazze Aperte exposure: binary

```{r}
#| echo: false
piazze_any_thematic_leaflet
```

### Piazze Aperte exposure: metres of affected street network

```{r}
#| echo: false
piazze_m_thematic_leaflet
```

### Piazze Aperte exposure: share of street network

```{r}
#| echo: false
piazze_share_thematic_leaflet
```
 
 