---
title: "Milan (2016–2022): active travel transformations (OSM vs official)"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
---

## Overview

Exploratory, visual comparison of active-travel infrastructure changes in Milan (2016–2022) using OpenStreetMap (OSM) and official municipal data.

Goal: support the Milan ATRAPA work by identifying which source (or combination) is most suitable for measuring active travel intervention changes over time.

## Data sources

### OpenStreetMap (OSM)

Historical OSM snapshots (160101, 220101) downloaded with `osmextract::oe_get()` and clipped to the Milan perimeter. OSM categories used in this workflow:

- Cycling infrastructure: `strong_ci` (separated), `moderate_ci` (painted lane), `weak_ci` (mixed traffic), `shared_foot` (shared with pedestrians)

- Pedestrian-priority: `pedestrian_street`, `living_street`

### Official municipal data (Milan)

From the Comune di Milano open data portal ([open access](https://dati.comune.milano.it/organization/da720f96-c055-4abb-bcce-d26eb8cf436a?groups=tran)), cleaned and harmonised for a 2016–2022 comparison:

- Cycling infrastructure (lines): typology (ciclabile sede propria; ciclabile segnaletica; promiscuo pedonali; promiscuo veicoli)
- Pedestrian areas + ZTL (polygons): classified by `tipo` (AP pedestrian area; ZTL limited traffic zone; AREA_B/AREA_C Area B/C; CR reserved carriageway), with activation date in `val_inizio` when available


## What is mapped

- 2016 network
- Added/removed segments (2016–2022)
- 2022 network
- Segment-level inspection via Google Street View (GSV) links

<!-- ## Run the workflow -->

<!-- Scripts used: -->
<!-- - `R/00_setup.R` -->
<!-- - `R/01_get_boundary.R` -->
<!-- - `R/02_osm_data.R` -->
<!-- - `R/03_official_data.R` -->
<!-- - `R/04_differences.R` -->
<!-- - `R/05_maps.R` -->

## Interactive maps

::: {.panel-tabset}

### Cycling infrastructure

```{r}
#| message: false
#| warning: false
#| include: false

source("R/00_setup.R")
source("R/01_get_boundary.R")
source("R/02_osm_data.R")
source("R/03_official_data.R")
source("R/04_differences.R")
source("R/05_maps.R")
```

```{r}
#| echo: false
#| warning: false
#| message: false

make_six_maps_cycling()
```

### Pedestrian-priority infrastructure

```{r}
#| echo: false
#| message: false
#| warning: false
make_six_maps_ped_priority()
```

:::